---
title: "Deciding the number of clusters using FractCurve package."
author: Shimpei Morimoto
output: 
  html_document:
    mathjax:  default
    fig_caption:  true
    toc: true
    section_numbering: true
vignette: >
  %\VignetteIndexEntry{Deciding the number of clusters using FractCurve package.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
	fig.width = 12, 
	message = FALSE,
	warning = FALSE,
	tidy.opts = list(
		keep.blank.line = TRUE,
		width.cutoff = 150
		),
  
	options(width = 150),
	eval = TRUE
)
library(FractCurve)
```

# 1. Introduction.
`fract_curve()` function in __*FractCurve*__ package can be used for various purpose. Deciding the number of clusters in an exploratory data analyses is one of them. In this vignette, an explanation of the algorithm for utilizing `fract_curve()` function in this purpose is presented. Also, a result from simulation study is shown. The analysis can be conducted from the `fract_curve_clust()` function, which is a wrapper function of `fract_curve()`, in this package. If you have not read the vignette on `fract_curve()`, you can invoke from this package by using `vignette("FractCurve", "FractCurve")`.  


# 2. A result from simulation.
Suppose that you you have a dataset with the size of 50, the features of 32 variables. There are 8 subgroups in the population from which the data was sampled, and the characteristics of the subgroups are represented by the 32 features. This situation is simulated by the `R` code below:

```{r simulation}
for(i in 1:10){
mat_i <- matrix(
  c(
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1), 
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1), 
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1),
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1), 
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1), 
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1),
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1),
    rnorm(200,rnbinom(n = 1, size = 10, mu = 5)-1,1)
    ),
  ncol = 50
  )
if(i==1){mat <- mat_i}else{mat <- rbind(mat,mat_i)}
}

heatmap(mat)
```

Using `fract_curve_clust()` function, that is a wrapper function of `fract_curve()`, the number of subgroups in the population can be detected as below.

```{r a result from the simulation}

res.fract_curve_clust <- fract_curve_clust(
  df.features = as.data.frame(mat),
  method.dist.row = "manhattan", method.dist.col = "manhattan",
  method.hclust.row = "ward", method.hclust.col = "ward",
  dir.output = NULL,
  get.df_of_IYs=FALSE,
  fn.plot_pdf =NULL,
  fn.df_of_IYs = NULL,
  fisher_test = FALSE
  )
```



# 3. The algorithm.
we are seeking.


# 4 The two conditions to be met for the algorithm to work.
As explained in the previous section, .



